---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# GeneNeighborhood

The goal of GeneNeighborhood is to extract and analyze the orientation and proximity of upstream/downstream genes.

## Installation

(NOT THERE YET!) You can install the released version of GeneNeighborhood from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("GeneNeighborhood")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("pgpmartin/GeneNeighborhood")
```
## Example

Load the library:
```{r load libraries}
library(GeneNeighborhood)
```

```{r load hidden libs, include = FALSE}
library(knitr)
library(dplyr)
```
  

### Obtain data on the gene neighbors

To run examples, the package includes a *GRanges* named **Genegr** that contains 676 genes with random coordinates on a single chromosome *Chr1*:
```{r gr data}
Genegr
```
  

For each feature/gene, we extract information (orientation and distance, potential overlaps) about their upstream/downstream neighbors with:
```{r extract GeneNeighbors}
GeneNeighbors <- getGeneNeighborhood(Genegr)
```
  

### Analyze the orientation of the genes' neighbors
  
We define a random set of 100 genes:
```{r define random genes}
set.seed(123)
randGenes <- names(Genegr)[sample.int(676, 100)]
```
  
  
We extract statistics about the orientation of their neighbors using:
```{r NOS random genes}
## Neighbors Orientation Statistics:
NOS <- analyzeNeighborsOrientation(randGenes, 
                                   GeneNeighborhood = GeneNeighbors)
```
  
  
By default all genes are used as a universe and an enrichment test is performed.   
By default, the function also analyzes the *"other"* orientation which may be hard to interpret. We can remove this orientation using:
```{r NOS without other}
NOS <- analyzeNeighborsOrientation(randGenes, 
                                   GeneNeighborhood = GeneNeighbors,
                                   keepOther = FALSE)
```
  
  
We obtain the following table:
```{r NOS kable, echo = FALSE}
NOS  %>% 
  dplyr::mutate(p.value = signif(p.value, 2)) %>%
  knitr::kable(digits = c(1,1,1,2,1,2,300), 
               format = "markdown",
               caption="100 random genes")
```
  
  
We can plot the corresponding percentages using:
```{r plotOrientationNOS}
plotNeighborsOrientation(NOS)
```


### Analyze the proximity of the genes' neighbors

We can analyze specifically the distances to the upstream genes with:
```{r}
randUpstreamDist <- statDistanceSide(GeneNeighborhood = GeneNeighbors,
                                     glist = randGenes,
                                     Side = "Upstream")
```

Which gives the following (simplified) table:
```{r randUpstreamDist kable, echo = FALSE}
randUpstreamDist$stats  %>% 
  dplyr::select(GeneGroup, SideClass, n, Median, Mean, SD, KS.pvalue:Independ.pvalue) %>%
  dplyr::mutate(KS.pvalue = signif(KS.pvalue, 2),
                Wilcox.pvalue = signif(Wilcox.pvalue, 2),
                Independ.pvalue = signif(Independ.pvalue, 2)) %>%
  knitr::kable(digits = c(1,1,1,2,2,2,300,300,300), 
               format = "markdown",
               caption="Upstream distances")
```
  
  
Or we directly analyze the distances to both upstream and downstream genes:
```{r}
randDist <- analyzeNeighborsDistance(GeneList = randGenes,
                                     GeneNeighborhood = GeneNeighbors)
```

Which gives the following (simplified) table:
```{r randDist kable, echo = FALSE}
randDist$stats  %>% 
  dplyr::select(GeneGroup:n, Median, Mean, SD, KS.pvalue:Independ.pvalue) %>%
  dplyr::mutate(KS.pvalue = signif(KS.pvalue, 2),
                Wilcox.pvalue = signif(Wilcox.pvalue, 2),
                Independ.pvalue = signif(Independ.pvalue, 2)) %>%
  knitr::kable(digits = c(1,1,1,1,2,2,2,300,300,300), 
               format = "markdown",
               caption="Upstream distances")
```

