---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# GeneNeighborhood

The goal of GeneNeighborhood is to extract and analyze the orientation and proximity of upstream/downstream genes.

## Installation

You can install the released version of GeneNeighborhood from [CRAN](https://CRAN.R-project.org) with (NOT THERE YET!):

``` r
install.packages("GeneNeighborhood")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("pgpmartin/GeneNeighborhood")
```
## Example

Load the library:
```{r}
library(GeneNeighborhood)
```
  

To run examples, the package includes a *GRanges* named gr that contains 676 genes with random coordinates on a single chromosome *Chr1*:
```{r gr data}
gr
```
  
  
For each feature/gene, we extract information (orientation and distance, potential overlaps) about their upstream/downstream neighbors with:
```{r extract GeneNeighbors}
GeneNeighbors <- GetGeneNeighborhood(gr)
```


